---
title: "Two Sample Ttest Power Analysis"
output:
  rmarkdown::github_document: default
  github_notebook: default
---

```{r echo=FALSE, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dpi=700, fig.cap=TRUE)
```

### Use case framing
- 

### Why a power analysis?
- Power analysis can help us determine the optimal sample size per group for an experiment.
- We're interested in deriving the power of an expected or observed set of experiment data.

### What do the various inputs represent?
- Two sample t test power analysis uses the following inputs.
- Power, Sig Level, Sample Size, N 

- Sig level: probability of observing an effect by chance
- Power: probability of detecting an effect when one exists
- Effect: 
- Delta:
- Standard deviation:

```{r}
library(tidyverse)
library(infer)
library(ggrepel)
library(patchwork)
```

```{r}
ttest_target_sample_size_fun <- function(delta_var, sd_var) {
      res <- power.t.test(delta = delta_var, sd = sd_var, 
                   sig.level = 0.05, power = 0.8, 
                   type = "two.sample")$n
      ceiling(res)
}

ttest_target_sample_size_fun(delta_var = 0.75, sd_var = 2.25)

expand_grid(delta_input = seq(1, 5, 0.5),
            sd_input = seq(0.25, 5, 0.25)) %>%
      rowwise() %>%
      mutate(samples_per_group = ttest_target_sample_size_fun(
                  delta_var =delta_input, sd_var = sd_input
            )
      ) %>%
      ggplot(aes(x=sd_input,
                 y=samples_per_group,
                 group=factor(delta_input),
                 color=factor(delta_input))) +
      geom_point() +
      geom_line()
```
### Related topics / further research
- Variance reduction techniques which can help speed up A/B tests and/or increase statistical power for a test
-



